``` {r setup, echo=FALSE, message=FALSE, include=FALSE, error=FALSE}
library(dplyr)

# Output folder for this document
options(knitr.figure_dir = "calculate_enrichments_at_known_enhancers")

source("shared_code/knitr_common.r")
source("shared_code/granges_common.r")
source("shared_code/enhancers.r")
source("shared_code/samples.r")
```

# Calculate histone and transcription factor enrichments at known enhancers

``` {r header_child, child="child_docs/header_child.Rmd"}
```

## Overview

For each known enhancer, we will calculate an enrichment value for all histone marks and transcription factors in each tissue.

For transcription factors, the enrichment will be calculated using a 201bp window centered at the summit within the enhancer.

For histone marks, the enrichment will be calculated using a 1001bp window centered on the enhancer itself.

The whole cell extract samples used for enrichment calculations are:

  * Toll10b: `tl10b_combined_wce` (all available Toll10b WCE samples merged into one)
  * gd7: `gd7_combined_wce` (all available gd7 WCE samples merged into one)

Enrichment is calculated using the following equation, 

$$enrichment = \frac{\frac{ChIP\ signal\ in\ region}{total\ ChIP\ signal}}{\frac{WCE\ signal\ in\ region}{total\ WCE\ signal}}$$

``` {r calculate_enrichments}

single_sample_enrichment <- function(i, samples.df, enhancers.gr) {
  stopifnot(length(i) == 1)

  mcols(enhancers.gr) <- mcols(enhancers.gr)[, c("enhancer_name"), drop=FALSE]
  enhancers.gr$enhancer
  enhancers.gr$sample_name <- samples.df$sample[i]
  enhancers.gr$tissue <- samples.df$tissue[i]
  enhancers.gr$factor <- samples.df$factor[i]
  enhancers.gr$factor_type <- samples.df$type[i]
  enhancers.gr$replicate <- samples.df$replicate[i]
  
  regions.gr <- enhancers.gr
  
  ip.bigwig <- sample_bigwig(samples.df$sample[i], "ip")
  wce.bigwig <- sample_bigwig(samples.df$sample[i], "wce")
  
  if(samples.df$type[i] == "histone") {
    regions.gr <- resize(regions.gr, width=1001, fix="center")
  } else {
    start(regions.gr) <- end(regions.gr) <- regionWhichMaxs(enhancers.gr, ip.bigwig)
    regions.gr <- resize(regions.gr, width=201, fix="center")
  }
  
  enhancers.gr$ip_signal <- regionSums(regions.gr, ip.bigwig)
  enhancers.gr$wce_signal <- regionSums(regions.gr, wce.bigwig)
  enhancers.gr$total_ip_signal <- total_signal(ip.bigwig)
  enhancers.gr$total_wce_signal <- total_signal(wce.bigwig)
  enhancers.gr$enrichment <- with(mcols(enhancers.gr), (ip_signal / total_ip_signal) / (wce_signal / total_wce_signal))
  as.data.frame(enhancers.gr)
}

enrichments.df <- cache("enrichments.df", function() {
  1:nrow(samples.df) %>%
    checked_mclapply(single_sample_enrichment, samples.df, enhancers.gr, mc.cores=6, mc.preschedule=FALSE) %>%
    bind_rows
})

```

``` {r session_info_child, child="child_docs/session_info_child.Rmd"}
```
