``` {r setup, echo=FALSE, message=FALSE, include=FALSE, error=FALSE}
library(ggplot2)
library(dplyr)

# Output folder for this document
options(knitr.figure_dir = "figure_1b_transcription_factor_boxplots")

source("shared_code/knitr_common.r")
source("shared_code/ggplot_theme.r")
source("shared_code/granges_common.r")
source("shared_code/samples.r")
source("shared_code/flybase.r")
source("shared_code/enhancers.r")
source("shared_code/import.r")
source("shared_code/dhs.r")
```

# Figure 1B: Transcription factor boxplots

``` {r header_child, child="child_docs/header_child.Rmd"}
```

``` {r load_data}
enrichments.df <- readRDS("calculate_enrichments_at_known_enhancers/enrichments.df.rds")
dhs.grl <- load_dhs()
late_k27ac.gr <- import.modencode_gff("data/external/modencode/4120_repset.17400896.enrichment.clusters.gff3.gz")
```

## Closed regions

Closed regions are putative enhancer regions active in the late (14-16h AEL) embryo but not active in the early embryo (2-4h AEL). These are defined by the following criteria:

  * Defined as open by DHS only in Stage 14 but no previous stages
  * More than 1,000 bp from an annotated TSS
  * Enriched in H3K27Ac in 14-16h (using modENCODE's list of enriched regions)
  * At least 200 bp in width

``` {r define_dhs_regions}

dhs.14 <- dhs.grl[["S14"]]
dhs.other <- dhs.grl[names(dhs.grl) != "S14"] %>% 
             unname %>%
             do.call(c, .) %>%
             reduce %>%
             coverage

dhs.14$early_score <- regionSums(dhs.14, dhs.other)
late_dhs <- dhs.14[dhs.14$early_score == 0]

late_dhs <- intersect(late_dhs, late_k27ac.gr)

late_dhs <- late_dhs[width(late_dhs) >= 200]

set.seed(12345)

late_dhs %<>% assign_nearest_gene %>%
              subset(distance_to_tss > 1000) %>%
              sample(100)

saveRDS(late_dhs, file=figure_path("late_dhs_with_k27ac.gr.rds"))
```

``` {r select_top_replicate}

factors <- c("dl", "twi", "mad", "zen", "zld")

medians.df <- enrichments.df %>%
              filter(factor %in% factors) %>%
              group_by(tissue, factor, replicate) %>%
              summarize(median_enrichment = median(enrichment))

replicates.df <- medians.df %>%
                 group_by(tissue, factor) %>%
                 summarize(best_median_enrichment = max(median_enrichment),
                           replicate              = replicate[which.max(median_enrichment)])

medians.df %>% pander(caption="Median enrichment values")
replicates.df %>% pander(caption="Replicate selection")
```

``` {r add_enrichments_to_dhs_regions}

find_bigwig <- function(factor, tissue, replicate, type="ip") {
  sample_name <- samples.df$sample[samples.df$factor == factor & samples.df$tissue == tissue & samples.df$replicate == replicate]
  sample_bigwig(sample_name, type)
}

add_enrichments <- function(gr, factor, tissue, replicate) {
  ip.bw <- find_bigwig(factor, tissue, replicate, "ip")
  wce.bw <- find_bigwig(factor, tissue, replicate, "wce")
  
  regions.gr <- gr
  start(regions.gr) <- end(regions.gr) <- regionWhichMaxs(regions.gr, ip.bw)
  regions.gr %<>% resize(201, fix="center")
  
  gr$ip_signal <- regionSums(regions.gr, ip.bw)
  gr$wce_signal <- regionSums(regions.gr, wce.bw)
  gr$total_ip_signal <- total_signal(ip.bw)
  gr$total_wce_signal <- total_signal(wce.bw)
  gr$enrichment <- with(mcols(gr), (ip_signal / total_ip_signal) / (wce_signal / total_wce_signal))
  gr$tissue <- tissue
  gr$factor <- factor
  gr$replicate <- replicate
  gr
}

dhs_enrichments.df <- 1:nrow(replicates.df) %>%
                      lapply(function(i) {
                        i.gr <- add_enrichments(late_dhs, replicates.df$factor[i], replicates.df$tissue[i], replicates.df$replicate[i])
                        as.data.frame(i.gr)
                      }) %>%
                      bind_rows

dhs_zelda.df <- subset(dhs_enrichments.df, factor == "zld")
dhs_zelda.df$factor <- "zld (both)"
dhs_zelda.df$group <- "Closed regions"

dhs_enrichments.df$factor[dhs_enrichments.df$factor == "zld" & dhs_enrichments.df$tissue == "de"] <- "zld_de"
dhs_enrichments.df$factor[dhs_enrichments.df$factor == "zld" & dhs_enrichments.df$tissue == "m"] <- "zld_m"

dhs_enrichments.df %<>% select(enrichment, factor)

dhs_enrichments.df$group <- "Closed regions"

m_enhancers  <- subset(enhancers.df, activity == "m")$enhancer_name
de_enhancers <- subset(enhancers.df, activity == "de")$enhancer_name

active_group <- bind_rows(enrichment_values("m",  "twi", "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers),
                          enrichment_values("m",  "dl",  "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers),
                          enrichment_values("m",  "zld", "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers),
                          enrichment_values("de", "mad", "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers),
                          enrichment_values("de", "zen", "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers),
                          enrichment_values("de", "zld", "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers))

active_zelda <- subset(active_group, factor == "zld")
active_zelda$group <- "Active enhancers"

active_group$factor[active_group$factor == "zld" & active_group$tissue == "de"] <- "zld_de"
active_group$factor[active_group$factor == "zld" & active_group$tissue == "m"] <- "zld_m"
active_group$group <- "Active enhancers"

repressed_group <- bind_rows(enrichment_values("m",  "twi", "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers),
                             enrichment_values("m",  "dl",  "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers),
                             enrichment_values("m",  "zld", "best", enrichments.df) %>% filter(enhancer_name %in% de_enhancers),
                             enrichment_values("de", "mad", "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers),
                             enrichment_values("de", "zen", "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers),
                             enrichment_values("de", "zld", "best", enrichments.df) %>% filter(enhancer_name %in% m_enhancers))

repressed_zelda <- subset(repressed_group, factor == "zld")
repressed_zelda$group <- "Repressed enhancers"

repressed_group$factor[repressed_group$factor == "zld" & repressed_group$tissue == "de"] <- "zld_de"
repressed_group$factor[repressed_group$factor == "zld" & repressed_group$tissue == "m"] <- "zld_m"
repressed_group$group <- "Repressed enhancers"
```

## Enrichments figure

``` {r build_plot}

all_zelda <- rbind(active_zelda, repressed_zelda)
all_zelda$factor <- "zld (both)"

cols <- c("factor", "enrichment", "group")

groups.df <- bind_rows(active_group[, cols],
                       repressed_group[, cols],
                       all_zelda[, cols],
                       dhs_zelda.df[, cols],
                       dhs_enrichments.df[, cols])

groups.df$group <- factor(groups.df$group, levels=c("Active enhancers", "Repressed enhancers", "Closed regions"))

group_colors <- c("Active enhancers"="#EB423D", "Repressed enhancers"="#B62324", "Closed regions"="#631A29")

groups.df$factor <- factor(groups.df$factor, levels=c("dl", "twi", "zld_m", "mad", "zen", "zld_de", "zld (both)"))

g <- ggplot(groups.df, aes(x=factor, y=log2(enrichment), fill=group)) +
     geom_boxplot(outlier.size=1) +
     scale_fill_manual("", values=group_colors) +
     labs(x="", y="Normalized ChIP-seq\nenrichments over input (log2)\n") +
     theme_manuscript()

```

``` {r display_plot, fig.width=9, fig.height=5.5, dev=c("png", "pdf")}
g
```

### Statistics

``` {r statistics}

wilcox_pvalue_by_group <- function(enrichments, groups, group1, group2) {
  wilcox.test(x = enrichments[groups == group1],
              y = enrichments[groups == group2],
              alternative = "two.sided",
              paired = FALSE)$p.value
}

stats.df <- groups.df %>%
            group_by(factor) %>%
            summarize(active_vs_repressed_pvalue = wilcox_pvalue_by_group(enrichment, group, "Active enhancers", "Repressed enhancers"),
                      repressed_vs_closed_pvalue = wilcox_pvalue_by_group(enrichment, group, "Repressed enhancers", "Closed regions"),
                      active_n = sum(group == "Active enhancers"),
                      repressed_n = sum(group == "Repressed enhancers"),
                      closed_n = sum(group == "Closed regions"))

stats.df %>% pander
```

## Fold change figure

```{r zelda_fold_change}

enhancers <- enhancers.df %>%
             select(enhancer_name, gene_name, active_tissue = activity)

zld_de <- enrichment_values(tissue="de", factor="zld", replicate="best", enrichments.df) %>%
          mutate(zld.de = log2(enrichment)) %>%
          select(enhancer_name, zld.de)
              
zld_m  <- enrichment_values(tissue="m", factor="zld", replicate="best", enrichments.df) %>%
          mutate(zld.m = log2(enrichment)) %>%
          select(enhancer_name, zld.m) 

zelda_both <- merge(zld_de, zld_m) %>%
              merge(enhancers) %>%
              mutate(zld_fc = zld.m - zld.de)

```

```{r zelda_fold_change_plot, dev=c("png", "pdf"), fig.cap="", fig.width=3, fig.height=5}
zelda_both$active_tissue %<>% factor(levels=c("m", "de"))

g <- ggplot(zelda_both, aes(x=active_tissue, y=zld_fc, fill=active_tissue)) +
     geom_boxplot() +
     scale_fill_manual("", values=activity_colors) +
     theme_manuscript() +
     labs(x="", y="Fold change in Zelda\nToll/gd (log2)\n")
g
```

### Statistics

```{r zelda_fold_change_stats}

values_meso <- subset(zelda_both, active_tissue == "m")$zld_fc
values_de   <- subset(zelda_both, active_tissue == "de")$zld_fc

zelda_fc_test <- data_frame("Test"="Wilcox two-sided", n=nrow(zelda_both), pvalue=wilcox.test(x = values_meso, y = values_de, alternative = "two.sided", paired=FALSE)$p.value)

zelda_fc_test %>% pander(caption="Zelda fold change between meso and DE")
```

``` {r session_info_child, child="child_docs/session_info_child.Rmd"}
```
